---
title: "ptarmigan_formatting"
author: "Diana Bowler"
date: "2 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Decide whether we want to subset the data and whether we want to plot the lines/observations

```{r}

#myOmradeNavn = "Finnemarka"
#myFylkesnavn = c("Finnmark","Troms")
myOmradeNavn = NULL
myFylkesnavn = NULL
plot=T

```

First get the raw data

```{r cache=TRUE}

library(knitr)

knit("ptarmigan_readin.Rmd")

```


#working out what distance for the human observations to include
```{r}

mydataTemp<-subset(mydata,month%in%c(7,8,9))#anysurvey

#subset
dogObs<-subset(mydataTemp,FK_ArtId==1&FK_OppfluktId%in%c(1,2) & !is.na(LinjeAvstand) & LinjeAvstand<500)
humanObs<-subset(mydataTemp,FK_ArtId==1&FK_OppfluktId%in%c(3,4)& !is.na(LinjeAvstand) & LinjeAvstand<500)

#plotting
hist(dogObs$LinjeAvstand)
hist(humanObs$LinjeAvstand)

#Subset the human distances to have a close to half-normal distribution of distances
allDistances<-rbind(dogObs,subset(humanObs,LinjeAvstand<10))

#multiple the dataset by -1 and create a mirror
myDistances<-c(abs(allDistances$LinjeAvstand),(abs(allDistances$LinjeAvstand)*-1))
hist(myDistances)

#test a distribution is the same as a norma, one
ks.test(myDistances,pnorm)

hist(allDistances$LinjeAvstand)

```


#Subset the data for the moment to willow ptarmigan, found by a dog, in July, August or September

```{r}

mydataTemp<-subset(mydata,month%in%c(7,8,9))#anysurvey
dataSubset1<-subset(mydataTemp,FK_ArtId==1 & FK_OppfluktId%in%c(1,2))#willow partmigan seen and any data flushed by a dog
dataSubset2<-subset(mydataTemp,FK_ArtId==1 & FK_OppfluktId%in%c(3,4) & abs(LinjeAvstand)<10)#willow partmigan seen and any data flushed by a dog
dataSubset<- rbind(dataSubset1,dataSubset2)
dataSubset$LinjeAvstand <- abs(dataSubset$LinjeAvstand)#get rid of negative values
detach("package:plyr")

```


Selection decision:

#(1)Focus on the sites recorded since 2007 onwards and also remove Nordland

```{r}
#get list of lines surveys 

library(tidyverse)

tempDF<-dataSubset %>%
  filter(Year==2007) %>%
    group_by(OmradeNavn) %>%
      summarise(n=length(Adults))
tempDF

dataSubsetF<-dataSubset %>%
  filter(Fylkesnavn %in% c("Finnmark","Troms"))

```

This dataset just contains data from: 
Finnmark             
Troms 

#Plot the data again
```{r,results='asis'}

#get total birds per transect
dataSubsetSummmary <- dataSubsetF %>%
    group_by(Latitude,Longitude,Year) %>%
      summarise(totalAdults = sum(Adults))


ggplot(subset(dataSubsetSummmary, Latitude > 64 & Latitude < 75))+
  geom_point(aes(x=Longitude,y=Latitude,colour=log(totalAdults)))+
  facet_wrap(~Year)+
  scale_colour_gradient(low="blue",high="red")

```

#Lets look at a time series for each line

```{r}

dataSubsetSummmary <- 
  dataSubsetF %>%
    group_by(LinjeID,Year,Kommunenavn) %>%
      summarise(totalAdults = sum(Adults))

ggplot(dataSubsetSummmary)+
  geom_line(aes(x=Year,y=totalAdults,colour=factor(LinjeID)))+
  facet_wrap(~Kommunenavn)+
  theme(legend.position = "none")

```

#(2) Or rather subset by well-sampled lines

```{r}
library(tidyverse)

tempDF<-dataSubset %>%
    group_by(LinjeID) %>%
      summarise(n=length(unique(Year[!is.na(totalIndiv)])))
tempDF

dataSubset<-dataSubset %>%
  filter(LinjeID %in% tempDF$LinjeID[tempDF$n>5])#Lines sampled in at least 5 years

```


#Plot the data again
```{r results='asis'}

#get total birds per transect
dataSubsetSummmary <- dataSubset %>%
    group_by(Latitude,Longitude,Year) %>%
      summarise(totalAdults = sum(Adults))


ggplot(subset(dataSubsetSummmary,Latitude>50 &Latitude<75))+
  geom_point(aes(x=Longitude,y=Latitude,colour=log(totalAdults)))+
  facet_wrap(~Year)+
  scale_colour_gradient(low="blue",high="red")

```

#Lets look at a time series for each line

```{r results='asis'}

dataSubsetSummmary <- 
  dataSubset %>%
    group_by(LinjeID,Year,Rapporteringsniva) %>%
      summarise(totalAdults = sum(Adults))

ggplot(subset(dataSubsetSummmary,Year>2006))+
  geom_line(aes(x=Year,y=totalAdults,colour=factor(LinjeID)))+
  facet_wrap(~Rapporteringsniva)+
  theme(legend.position = "none")+
  scale_y_log10()

```

#Lets fit a distance model again

```{r cache = TRUE}


#truncate distances at 200
dataSubset = dataSubset %>%
        filter(LinjeAvstand<200,!is.na(LinjeAvstand))

dataSubset$LinjeAvstand <- abs(dataSubset$LinjeAvstand)

nrow(dataSubset)#4678

#bound extremes of timeSince

dataQuantile <- quantile(abs(dataSubset$timeSince),0.95,na.rm=T)#360 - 6 hours..., probably longer than this is a new survey???
dataSubset$timeSince<-ifelse(abs(dataSubset$timeSince)>dataQuantile, median(abs(dataSubset$timeSince)), abs(dataSubset$timeSince))

#examine the group sizes
dataSubset$AntallUkjent[dataSubset$AntallUkjent==210]<-0
dataSubset$totalIndiv<-dataSubset$Adults+dataSubset$AntallKylling+dataSubset$AntallUkjent
hist(dataSubset$totalIndiv)

#organise data for a distance model
bugs.data <- list(W = 200,
            N = nrow(dataSubset),
            y = dataSubset$LinjeAvstand,
            GroupSize = log(dataSubset$totalIndiv+1),
            #GroupSize2 = (log(dataSubset$totalIndiv+1))^2,
            Time = log(dataSubset$timeSince+1),
            zeros.dist = rep(0,nrow(dataSubset)))

source('C:/Users/diana.bowler/OneDrive - NINA/methods/models/bugsFunctions.R')

setwd("C:/Users/diana.bowler/OneDrive - NINA/Alpine/Ptarmigan/Scripts")

params <- c("b.group.size","b.time","esw")
  
#inits <- function(){list(b.df.0 = runif(1,2,5), 
#                         B.gs.0 = runif(1,0.2,3),
#                         B.n.0 = runif(1,0.5,5))}
  
#out1 <- jags(bugs.data, inits=NULL, params, "ptarmigan_distance_model.txt", n.thin=nt,
#               n.chains=nc, n.burnin=nb,n.iter=ni)

```

#Look at summary ouput from this model

```{r}
#out1$summary[grepl("esw",row.names(out1$summary)),]
#summary(out1$mean$esw)

#print(out1,3)
#Size effect but not time effect

```

#Put the zeros and NAs in the dataset

```{r}
library(plyr)

#First insert the 0s into the datasheet
lineInfo <- unique(mydataTemp[,c("Year","LinjeID","Fylkesnavn","Kommunenavn","OmradeNavn","OmradeID","Rapporteringsniva","TakseringID","Date","Year","month","Time","startTime","timeSince","LengdeTaksert")])
#use Rapport instead of Fylk/Kommune??
allData <- merge(lineInfo,dataSubset,all.x=T,by=c("Year","LinjeID","Fylkesnavn","Kommunenavn","OmradeNavn","OmradeID","Rapporteringsniva","TakseringID","Date","Year","month","Time","startTime","timeSince","LengdeTaksert"))
#allData[,c("AntallHann","AntallHunn","AntallKylling","AntallUkjent","totalIndiv","Adults")])<-0#why wont work??
allData$AntallHann[is.na(allData$AntallHann)]<-0
allData$AntallHunn[is.na(allData$AntallHunn)]<-0
allData$AntallKylling[is.na(allData$AntallKylling)]<-0
allData$AntallUkjent[is.na(allData$AntallUkjent)]<-0
allData$totalIndiv[is.na(allData$totalIndiv)]<-0
allData$Adults[is.na(allData$Adults)]<-0

#subset(lineInfo,LinjeID=="816"&Year==2001) - missing transect length data
#we are missing many other transect lengths...

#now add NAs when a site was not visited in a year
mydataWillow<-subset(allData,totalIndiv>0)
lineInfo <- unique(mydataWillow[,c("LinjeID","Fylkesnavn","Kommunenavn","OmradeNavn","OmradeID","Rapporteringsniva")])
potentialSurveys <- ldply(unique(mydataWillow$Year),function(x)cbind(Year=x,lineInfo))
allData <- merge(potentialSurveys,allData,by=c("Year","LinjeID","Fylkesnavn","Kommunenavn","OmradeNavn","OmradeID","Rapporteringsniva"),all.x=T)

#now restrict to lines in which a willow ptarmigan was seen at least once
sumStats<-ddply(allData,.(LinjeID),summarise,nuObs=sum(totalIndiv,na.rm=T))
allData <- subset(allData,LinjeID%in%sumStats$LinjeID[sumStats$nuObs>0])

#trim white space
allData$Fylkesnavn<-trim(as.character(allData$Fylkesnavn))

summary(allData$totalIndiv)

```
```{r}

#Explore the spatial structure

spatialCluster<-unique(allData[,c("LinjeID","Fylkesnavn","Kommunenavn","OmradeNavn","Rapporteringsniva")])

#first 3 are nested hierarchically

#Rapporteringsniva is between OmradeNavn and Fylkesnavn??

#Fylkesnavn should definately be included
ddply(spatialCluster,.(Fylkesnavn),summarise,nuRap=length(unique(Rapporteringsniva)))
#then Rapporteringsniva
ddply(spatialCluster,.(Rapporteringsniva),summarise,nuOm=length(unique(OmradeNavn)))#most Rapporteringsniva have 1 o 2 OmradeNavn
#then Kommune
ddply(spatialCluster,.(Rapporteringsniva),summarise,nuKo=length(unique(Kommunenavn)))#most Rapporteringsniva have 1 or 2 Kommunenavn
#or OmradeNavn and Kommune instead of Rapporteringsniva
ddply(spatialCluster,.(OmradeNavn),summarise,nuKo=length(unique(Kommunenavn)))#most OmradeNavn have 1 Kommunenavn
#check line numbers
ddply(spatialCluster,.(OmradeNavn),summarise,nuLi=length(unique(LinjeID)))#most OmradeNavn have multiple lines
ddply(spatialCluster,.(Rapporteringsniva),summarise,nuLi=length(unique(LinjeID)))#most Rapporteringsniva have multiple lines

#choice is between Rapporteringsniva and OmradeNavn
#either way  - dont use Kommune
#either way - use Fylkesnavn

length(unique(spatialCluster$Fylkesnavn))#10
length(unique(spatialCluster$OmradeNavn))#81
length(unique(spatialCluster$Rapporteringsniva))#40

```
