---
title: "Ptarmigan analysis"
author: "Diana Bowler"
date: "16 oktober 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```


##Access the database and retreive the data

Run the script to access the database and retreive the data

```{r, warning=FALSE, cache=TRUE}

source('C:/Users/diana.bowler/OneDrive - NINA/Alpine/Ptarmigan/Scripts/HFP-connection_DB.R')

```


##Explore the data

The data frame is called mydata

Remember this includes all species at the moment...

```{r }

str(mydata)

```

#Explore whether our calculated distances deviate from the reported distances

```{r }

#N_dist = Distance to line estimated form reported positions
#avvik = Difference between reported "Linjeavstand" and the estimated N_dist

summary(mydata$avvik)

hist(mydata$avvik)

hist(mydata$LinjeAvstand)

hist(mydata$N_dist)

#the LinjeAvstand actually look more reasonable...

```


#Explore the counts

```{r }

hist(mydata$AntallHann)     
hist(mydata$AntallHunn)     
hist(mydata$AntallKylling)   
hist(mydata$AntallUkjent)  

#There are some extreme Ukjent counts (probably when an observer has seen a flock)
summary(mydata$AntallUkjent)
mydata$AntallUkjent[mydata$AntallUkjent>100]<-0

```


#Clean the time data

```{r}

#Date
library(lubridate)
mydata$Date <- as.character(mydata$Dato)
mydata$Date <- as.Date(mydata$Date,format="%Y-%m-%d")
mydata$Year <- year(mydata$Date)
mydata$month <- month(mydata$Date)
unique(mydata$Year)
unique(mydata$month)

#Time
mydata$StartKl <- as.character(mydata$StartKl)
mydata$SluttKl <- as.character(mydata$SluttKl)
mydata$Klokkeslett <- as.character(mydata$Klokkeslett)

#get minutes since midnight from Klokkeslett
clockS = function(t){
  if(grepl(":",t)){
  tsplit<-strsplit(t,":")
  as.numeric(tsplit[[1]][1])*60+as.numeric(tsplit[[1]][2])
  }
  else
    NA
}
mydata$Time <- sapply(mydata$Klokkeslett,clockS)
mydata$startTime <- sapply(mydata$StartKl,clockS)
mydata$timeSince <- mydata$Time - mydata$startTime
table(mydata$month)

#work out time since transect began being sampled

#diference between start time and klock time

```

#Look at the species and flushing methods

```{r}
table(mydata$FK_ArtId)

table(mydata$FK_OppfluktId)

```

#Look at the spatial clustering

```{r}

#Spatial variables

#TakseringID (observer)

#Linje_ID (line)

#Fylkesnavn (county)

#Kommunenavn (municipality)

#Rapporteringsniva (reporting)

#OmradeNavn (area)

tempDF <- unique(
  mydata[,c("TakseringID","LinjeID","Fylkesnavn","Kommunenavn","Rapporteringsniva","OmradeNavn")])
tempDF <- data.frame(lapply(tempDF,factor))

str(tempDF)

#trim the white spaces
library(gdata)
mydata$Fylkesnavn<-trim(mydata$Fylkesnavn)
mydata$Kommunenavn<-trim(mydata$Kommunenavn)

```


#Check all are aktiv?

```{r}

table(mydata$Aktiv)
#yes

#so there should be observations if it was not aktiv??
tempDF <- subset(mydata,Aktiv==0)
summary(tempDF$AntallHunn)

#there are observations when it was not Aktiv???

#no need to worry about this according to Erlend

```


#Estimate the effective strip width, on average

```{r,results='asis'}

#get the Distances
myDistances<-mydata$LinjeAvstand[!is.na(mydata$LinjeAvstand)]
myDistances<-abs(myDistances)
summary(myDistances)

#cut off 5% quantile
truncDistance<-quantile(myDistances,0.95,na.rm=T)#233m
myDistances[myDistances>200]<-NA
myDistances<-myDistances[!is.na(myDistances)]

library(Rdistance)
fit <- F.dfunc.estim(myDistances, likelihood="halfnorm")#effective strip width is 118 m
plot(fit)

```


#Example of the trends in abundance (just based on positive observations)

```{r results='asis'}

#plot maps for each year showing the abundance of adults

mydata$Adults <- mydata$AntallHann + mydata$AntallHunn
mydata$totalIndiv <- mydata$Adults + mydata$AntallUkjent + mydata$AntallKylling

#get total birds per transect
mydataSummmary <- 
  mydata %>%
    group_by(Latitude,Longitude,Year) %>%
      summarise(totalAdults = sum(Adults))


library(ggplot2)

ggplot(subset(mydataSummmary, Latitude > 50 & Latitude < 75))+
  geom_point(aes(x=Longitude,y=Latitude,colour=log(totalAdults)))+
  facet_wrap(~Year)+
  scale_colour_gradient(low="blue",high="red")

```














