---
title: "ptarmigan_trends_analysis"
author: "Diana Bowler"
date: "2 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Import the data - check the subsetting is as wanted

```{r}
#library(knitr)
#knit("ptarmigan_formatting.Rmd")
load("/data/home/diana.bowler/NorskPtarmigan/allData.RData")
head(allData)

```


Fitting the full model on total population - subset to 2007 onwards - including the distance model

Organise the data

```{r, warning=FALSE}

library(magrittr)
library(plyr); library(dplyr)

#subset
allData<-subset(allData,Year>2006)

#remove hyphens for help with subsetting
allData$Fylkesnavn<-gsub("-"," ",allData$Fylkesnavn)
allData$Rapporteringsniva<-gsub("-"," ",allData$Rapporteringsniva)

#change kommune name for Dovre Fjellstyre
allData$Kommunenavn[which(allData$Kommunenavn=="Dovre Fjellstyrene")]<-"Dovre"

#get number of individuals
nuIndivs <-
  allData%>%
  group_by(Year,LinjeID,Fylkesnavn,Kommunenavn,OmradeNavn,Rapporteringsniva, add = T) %>%
  summarise(nuAdults = sum(Adults), 
            nuJuvs = sum(AntallKylling),
            nuTotal = sum(totalIndiv),
            maxTransectLength=max(LengdeTaksert,na.rm=T))

nuIndivs$iYear <- nuIndivs$Year - min(nuIndivs$Year) + 1
summary(nuIndivs)

library(reshape2)
nuIndivs$Line<-paste(nuIndivs$Fylkesnavn,nuIndivs$Rapporteringsniva,sep="_")
nuIndivs$Line<-paste(nuIndivs$Line,nuIndivs$LinjeID,sep="-")
my.n <- acast(nuIndivs,Line~Year,value.var="nuTotal")

#get transect length data
allData$Line<-paste(allData$Fylkesnavn,allData$Rapporteringsniva,sep="_")
allData$Line<-paste(allData$Line,allData$LinjeID,sep="-")
transectLengths <- acast(allData,Line~Year,value.var="LengdeTaksert",fun=max,na.rm=T)
transectLengths[is.na(my.n)]<-0

#order site info
siteInfo<-data.frame(Line=row.names(my.n))
siteInfo$Fylkesnavn<-as.numeric(as.factor(sapply(siteInfo$Line,function(x)strsplit(as.character(x),"_")[[1]][1])))
siteInfo$Rapporteringsniva<-as.numeric(as.factor(sapply(siteInfo$Line,function(x)strsplit(as.character(x),"-")[[1]][1])))
siteInfo$LinjeID<-as.numeric(as.factor(siteInfo$Line))

#Create a data frame with only detection
allDetections<-subset(allData,totalIndiv>0&!is.na(totalIndiv))
allDetections<-arrange(allDetections,LinjeID)
head(allDetections)

#Organise fur bugs
bugs.data <- list(#For the state model
                  n.Lines = length(unique(nuIndivs$LinjeID)),
                  n.Years = length(unique(nuIndivs$Year)),
                  n.Sites = length(unique(nuIndivs$Fylkesnavn)),
                  n.Sites2 = length(unique(interaction(nuIndivs$Rapporteringsniva,nuIndivs$Fylkesnavn))),
                  n.LineYear = length(unique(nuIndivs$LinjeID))*length(unique(nuIndivs$Year)),
                  line = siteInfo$LinjeID,
                  site = siteInfo$Fylkesnavn,
                  site2 = siteInfo$Rapporteringsniva, 
                  year = (1:length(unique(nuIndivs$Year))),
                  NuIndivs = my.n,
                  TransectLength = transectLengths,
                  #For the distance model
                  W = 200,
                  N = nrow(allDetections),
                  y = allDetections$LinjeAvstand,
                  detectionGroupSize = log(allDetections$totalIndiv+1),
                  GroupSize = allDetections$totalIndiv,
                  zeros.dist = rep(0,nrow(allDetections)),
                  detectionLine =as.numeric(factor(allDetections$LinjeID)),
                  detectionYear = as.numeric(factor(allDetections$Year)),
                  detectionSite = as.numeric(factor(allDetections$Rapporteringsniva)),
                  detectionLineYear = as.numeric(factor(interaction(allDetections$LinjeID,allDetections$Year))))

```

Plot the data

```{r results='asis'}
library(ggplot2)

ggplot(nuIndivs,aes(x=Year,y=nuTotal,group=LinjeID))+
  geom_point(aes(colour=LinjeID))+
  geom_line(aes(colour=LinjeID))+
  facet_wrap(~Fylkesnavn)+
  theme(legend.position="none")

```
#look also at trends in the group size

```{r}

ggplot(subset(allData,!is.na(totalIndiv)),aes(x=Year,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=LinjeID))+
  geom_line(aes(colour=LinjeID))+
  facet_wrap(~Fylkesnavn)+
  theme(legend.position="none")

```

Run model in JAGS - with a long-term trend term

```{r}

source('C:/Users/diana.bowler/OneDrive - NINA/methods/models/bugsFunctions.R')

setwd("C:/Users/diana.bowler/OneDrive - NINA/Alpine/Ptarmigan/Scripts")

#inits <- function(){list(trend = dunif(-2,0))}

params <- c("int","beta.auto","line.d.sd","site.d.sd","year.d.sd","totNumbers","predESW")

#inits <- function(){list(b.df.0 = runif(1,2,5))}

out1 <- jags(bugs.data, inits=NULL, params, "combined_model_ar1.txt", n.thin=nt,
               n.chains=3, n.burnin=1000,n.iter=5000)
```

Print output

```{r results='asis'}

#library(ggmcmc)
#ggs_traceplot(ggs(out1$samples)) 

print(out1,2)

```

Fit the model without the trend term and just random year to year variation, that also varies by fylk and rapp

Run model in JAGS

```{r}
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

library(jagsUI)

params <- c("line.d.sd","year.d.sd","site.d.sd","site2.d.sd","syear.d.sd","s2year.d.sd","Density","meanNumbers")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_noTrend.txt", n.thin=nt,
               n.chains=3, n.burnin=1000,n.iter=5000,parallel = T)
```


```{r results='asis'}
#library(ggmcmc)
#ggs_traceplot(ggs(out1$samples)) 

print(out1,2)
save(out1,file="out1_ptarmigan_noTrends2.RData")

```

#print at a later stage
```{r}
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_ptarmigan_noTrends2.RData")
<<<<<<< HEAD
print(out1$summary,2)
=======
print(out1)
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893

```


#Pull out and plot the density estimates

```{r results='asis'}
#Have they converged
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_ptarmigan_noTrends2.RData")
densityDF<-data.frame(meanNu=out1$mean$Density)
densityDF$LinjeID<-1:nrow(densityDF)
densityDF<-merge(densityDF,siteInfo,by="LinjeID")
temp<-sapply(densityDF$Line,function(x)strsplit(as.character(x),"-")[[1]][1])
densityDF$Fylkesnavn<-sapply(temp,function(x)strsplit(x,"_")[[1]][1])
densityDF$Rapporteringsniva<-sapply(temp,function(x)strsplit(x,"_")[[1]][2])
densityDF$originalLinjeID<-sapply(densityDF$Line,function(x)strsplit(as.character(x),"-")[[1]][2])
densityDF<-melt(densityDF,id=c("LinjeID","Line","Fylkesnavn","Rapporteringsniva","originalLinjeID"))
densityDF$variable<-gsub("meanNu.","",densityDF$variable)
densityDF$Year<-as.numeric(as.character(densityDF$variable))+2006
densityDF$value<-as.numeric(as.character(densityDF$value))
qplot(Year,value,data=densityDF,group=LinjeID,colour=Rapporteringsniva,geom="path",facets=~Fylkesnavn)+
  theme_light()+
  theme(legend.position="none")+ylab("Predicted Density km2")
#each site seems to have its own dynamic over time as specified by the model
```

Look at spatial pattern
```{r}
#get average abundance per linjeID
linePredictions <- ddply(densityDF,.(originalLinjeID),summarise,density=mean(value))

#get line spatial info
setwd("/data/home/diana.bowler/climaticFiles")
load("Lines_utm.RData")

#add mean densities and remove those without mean density data
Lines_utm$density <- linePredictions$density[match(Lines_utm@data$LinjeID,linePredictions$originalLinjeID)]
Lines_utm <- subset(Lines_utm,!is.na(density))

#plot it
library(sp)
mySLDF_fortify <- fortify(Lines_utm)
mySLDF_fortify$density <- linePredictions$density[match(mySLDF_fortify$id,linePredictions$originalLinjeID)]

ggplot(mySLDF_fortify, aes(x=long, y=lat, group=group,colour=log(density+1))) + 
  geom_path(size=rel(3)) +
  theme_minimal()+
  scale_colour_gradient2(low="red",mid="gray90",high="dodgerblue",midpoint=2.51)+
  xlab("x")+ylab("y")

#get temperature map
setwd("/data/home/diana.bowler/GeoSpatialData/AdministrativeUnits/Norway_AdministrativeUnits/Processed")
counties<-readOGR(getwd(),layer="Norway_ADM_fylke_FLATE")
counties<-spTransform(counties,CRS(proj4string(Lines_utm)))
plot(counties)
counties@data$id <- 1:nrow(counties@data)
counties.fort <- fortify(counties, region='id')
ggplot(counties.fort, aes(x = long, y = lat, group = id)) + 
  geom_polygon(fill="white")+
  geom_path(data=mySLDF_fortify, aes(x=long, y=lat, group=group,colour=log(density+1)),size=rel(3)) +
  scale_colour_gradient2(low="red",mid="gray90",high="dodgerblue",midpoint=2.51)+
  xlab("x")+ylab("y")

```

