---
title: "combined_model_springOnset"
author: "Diana Bowler"
date: "14 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

#get the data on spring

```{r results='asis'}
setwd("/data/home/diana.bowler/climaticFiles")

load("springSnowData.RData")
springSnowData<-subset(springSnowData,site %in% allData$LinjeID)
springSnowData$onset[is.infinite(springSnowData$onset)]<-max(springSnowData$onset[is.finite(springSnowData$onset)])
springSnowData$OmradeID<-allData$OmradeID[match(springSnowData$site,allData$LinjeID)]
springSnowData$Fylkesnavn<-allData$Fylkesnavn[match(springSnowData$site,allData$LinjeID)]

#plotting
ggplot(data=springSnowData,aes(x=Year,y=onset,group=OmradeID))+
  geom_point(aes(colour=OmradeID))+
  geom_line(aes(colour=OmradeID))+
  theme(legend.position = "none")+
  facet_wrap(~Fylkesnavn)

```

#Organise the data

```{r}

myCovariate = springSnowData
myCovariate = subset(myCovariate,site%in%nuIndivs$LinjeID & Year%in%nuIndivs$Year)
covariateMatrix = my.n
rowNames<-as.character(sapply(row.names(my.n),function(x)strsplit(as.character(x),"#")[[1]][2]))
rowNames<-as.character(sapply(rowNames,function(x)strsplit(as.character(x),"-")[[1]][2]))

all(rowNames%in%myCovariate$site)

for(i in 1:nrow(covariateMatrix)){
  for(j in 1:ncol(covariateMatrix)){
    covariateMatrix[i,j] = myCovariate$onset[myCovariate$site==rowNames[i]&
                                              myCovariate$Year==dimnames(covariateMatrix)[[2]][j]]
  }
}

#split into spatial and temporal anomalies
spatialMatrix<-as.numeric(rowMeans(covariateMatrix))
temporalMatrix<-as.numeric(colMeans(covariateMatrix))

#make a DF
temporalDF<-data.frame(Year=dimnames(my.n)[[2]],value=temporalMatrix)
spatialDF<-data.frame(OmradeID=rowNames,value=spatialMatrix)

#look at summaries
summary(spatialMatrix)
summary(temporalMatrix)

#make anomalies
bugs.data$temporalMatrix <- apply(covariateMatrix,2,function(x)x-spatialMatrix)
bugs.data$spatialMatrix <- t(apply(covariateMatrix,1,function(x)x-temporalMatrix))
bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix^2#also add a polynomial term for the spatial variation

#plot the data

#temporal
tMmelted<-melt(bugs.data$temporalMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("temporal")

#spatial
tMmelted<-melt(bugs.data$spatialMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("spatial")

```

#Spatial pattern and temporal anomalies

the temporal anomalies are the same as above

the spatial pattern is now assumed to be fixed across years

```{r}

bugs.data$spatialMatrix <-spatialMatrix
bugs.data$spatialMatrix2 <-bugs.data$spatialMatrix^2
bugs.data$spatialMatrix<-bugs.data$spatialMatrix-median(bugs.data$spatialMatrix)
bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2-median(bugs.data$spatialMatrix2)

plot(bugs.data$spatialMatrix,
     apply(bugs.data$temporalMatrix,1,sd),main="spring onset")
```

Fit glm

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","site.d.sd",
            "b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT","predESW")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_glm.txt", n.thin=nt,
               n.chains=3, n.burnin=1000,n.iter=5000,parallel = T)

```

Print output

```{r}
print(out1,2)

``` 

Fit glm scaled

```{r}
bugs.data$spatialMatrix<-scale(bugs.data$spatialMatrix)

bugs.data$temporalMatrix<-scale(bugs.data$temporalMatrix)

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","site.d.sd",
            "b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_glm.txt", n.thin=nt,
               n.chains=3, n.burnin=500,n.iter=2000,parallel = T)
print(out1,2)

```

#Fit a basic model against abundance

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT","predESW")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=10000,parallel = T)

```


#plotting the results

```{r}

#effect of space
library(reshape2)
dataspatialMelt<-melt(out1$mean$pred.Space)
dataspatialMelt2<-melt(out1$q2.5$pred.Space)
dataspatialMelt3<-melt(out1$q97.5$pred.Space)
spatialMatrixMelt<-melt(bugs.data$spatialMatrix)
dataspatialMelt$springOnset<-spatialMatrixMelt$value
dataspatialMelt$lower<-dataspatialMelt2$value
dataspatialMelt$upper<-dataspatialMelt3$value

ggplot(dataspatialMelt)+geom_point(aes(x=springOnset,y=value))+geom_ribbon(aes(x=springOnset,ymin=lower,ymax=upper),alpha=0.5)+
  theme_classic()+ggtitle("Spatial variation")

#effect of time
library(reshape2)
dataspatialMelt<-melt(out1$mean$pred.Time)
dataspatialMelt2<-melt(out1$q2.5$pred.Time)
dataspatialMelt3<-melt(out1$q97.5$pred.Time)
spatialMatrixMelt<-melt(bugs.data$temporalMatrix)
dataspatialMelt$springOnset<-spatialMatrixMelt$value
dataspatialMelt$lower<-dataspatialMelt2$value
dataspatialMelt$upper<-dataspatialMelt3$value

ggplot(dataspatialMelt)+geom_point(aes(x=springOnset,y=value))+geom_ribbon(aes(x=springOnset,ymin=lower,ymax=upper),alpha=0.5)+
  theme_classic()+ggtitle("Temporal variation")


```

#Fit a model with ar1 term

```{r}

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto","beta.covariateS","beta.covariateT","pred.Time","pred.Space")

out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1.txt", n.thin=nt,
               n.chains=3, n.burnin=400,n.iter=5000,parallel = T)

```

Print the output

```{r}
print(out1,2)

save(out1,file="out1_springOnset.RData")

```

#plotting the results

```{r}

#effect of space
library(reshape2)
dataspatialMelt<-melt(out1$mean$pred.Space)
dataspatialMelt2<-melt(out1$q2.5$pred.Space)
dataspatialMelt3<-melt(out1$q97.5$pred.Space)
spatialMatrixMelt<-melt(bugs.data$spatialMatrix)
dataspatialMelt$springOnset<-spatialMatrixMelt$value
dataspatialMelt$lower<-dataspatialMelt2$value
dataspatialMelt$upper<-dataspatialMelt3$value

ggplot(dataspatialMelt)+geom_point(aes(x=springOnset,y=value))+geom_ribbon(aes(x=springOnset,ymin=lower,ymax=upper),alpha=0.5)+
  theme_classic()+ggtitle("Spatial variation")

#effect of time
library(reshape2)
dataspatialMelt<-melt(out1$mean$pred.Time)
dataspatialMelt2<-melt(out1$q2.5$pred.Time)
dataspatialMelt3<-melt(out1$q97.5$pred.Time)
spatialMatrixMelt<-melt(bugs.data$temporalMatrix)
dataspatialMelt$springOnset<-spatialMatrixMelt$value
dataspatialMelt$lower<-dataspatialMelt2$value
dataspatialMelt$upper<-dataspatialMelt3$value

ggplot(dataspatialMelt)+geom_point(aes(x=springOnset,y=value))+geom_ribbon(aes(x=springOnset,ymin=lower,ymax=upper),alpha=0.5)+
  theme_classic()+ggtitle("Temporal variation")


```

#specify climate as covariate 1

```{r}

bugs.data$spatialMatrix1 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix1 <- bugs.data$temporalMatrix

```

#specify rodents as climate 2

#also get the rodent data - run code in the rodent Rmd

```{r}

bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix2 <- bugs.data$temporalMatrix

qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix2,1,sd),main="Rodent sd against spring onset")+stat_smooth(method="lm")

```

```{r}

#scale
bugs.data$spatialMatrix1<-bugs.data$spatialMatrix1/sd(bugs.data$spatialMatrix1)
bugs.data$temporalMatrix1<-bugs.data$temporalMatrix1/sd(bugs.data$temporalMatrix1)
bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2/sd(bugs.data$spatialMatrix2)
bugs.data$temporalMatrix2<-bugs.data$temporalMatrix2/sd(bugs.data$temporalMatrix2)

#plot
hist(bugs.data$spatialMatrix1)
hist(as.numeric(bugs.data$temporalMatrix1))
hist(bugs.data$spatialMatrix2)
hist(as.numeric(bugs.data$temporalMatrix2))

```

interactive effects - between temporal rodent and spatial climate

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2",
            "beta.covariate_int",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springOnset(snow)_interaction.RData")

```

Assess model fit - look at intercept, line and autocorrelation effect

```{r, results='asis'}
library(ggmcmc)
ggs_traceplot(ggs(out1$samples,family="int.d"))
ggs_traceplot(ggs(out1$samples,family="beta.auto"))
ggs_traceplot(ggs(out1$samples,family="line.d.sd"))

```

Look at coeffs of interest
```{r, results='asis'}
library(ggmcmc)
ggs_traceplot(ggs(out1$samples,family="beta.covariateS_cov1"))
ggs_traceplot(ggs(out1$samples,family="beta.covariateS_cov2"))
ggs_traceplot(ggs(out1$samples,family="beta.covariateT_cov1"))
ggs_traceplot(ggs(out1$samples,family="beta.covariateT_cov2"))
ggs_traceplot(ggs(out1$samples,family="beta.covariate_int"))

```

```{r, results='asis'}

library(ggmcmc)
ggs_density(ggs(out1$samples,family="beta.covariateS_cov1"))
ggs_density(ggs(out1$samples,family="beta.covariateS_cov2"))
ggs_density(ggs(out1$samples,family="beta.covariateT_cov1"))
ggs_density(ggs(out1$samples,family="beta.covariateT_cov2"))
ggs_density(ggs(out1$samples,family="beta.covariate_int"))

```

Plot the coefficients

```{r}

modelSummary<-data.frame(out1$summary[grepl("beta.covariate",row.names(out1$summary)),])
modelSummary$parameter<-c("springOnset_spatial","rodents_spatial","springOnset_temporal","rodents_temporal","spatialSpring x temporalRodents")

ggplot(modelSummary)+
  geom_crossbar(aes(x=parameter,y=mean,ymin=X2.5.,ymax=X97.5.))+
  geom_hline(yintercept=0,color="red",linetype="dashed")+
  coord_flip()+
  theme_bw()

```

Look at predictions

```{r}

out2<-update(out1,parameters.to.save="expNuIndivs.new",n.iter=2000)
ggs_density(ggs(out2$samples,family="expNuIndivs.new"))

my_ggs_ppmean(ggs(out1$samples,family="expNuIndivs.new"),outcome=as.numeric(bugs.data$NuIndivs))

```

Dynamics predicted by the model

```{r}
out2<-update(out1,parameters.to.save="Density",n.iter=2000)

modelSummary<-getBUGSFits(out2)

ggplot(subset(modelSummary,Year>2007),
       aes(x=Year,y=mean,group=Line))+
  geom_point(aes(colour=Kommunenavn))+
  geom_line(aes(colour=Kommunenavn))+
  facet_wrap(~Fylkesnavn)+
  theme(legend.position="none")+
  theme_bw()

```

Get model predictions with and without the rodent/climate terms

```{r}



```


interactive effects - between temporal rodent and spatial climate - incl lagged effects

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2","beta.covariateTL_cov2",
            "beta.covariate_int","beta.covariate_intL",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction_lagged.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springOnset(snow)_interaction_lagged.RData")

```

Weather interaction

scaled units

```{r}

bugs.data$spatialMatrix<-bugs.data$spatialMatrix/sd(bugs.data$spatialMatrix)

bugs.data$temporalMatrix<-bugs.data$temporalMatrix/sd(bugs.data$temporalMatrix)

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT",
            "beta.covariate_int",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_weather_interaction.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springOnset(snow)_weather_interaction_scaled.RData")

```

```{r}

modelSummary<-data.frame(out1$summary[grepl("beta.covariate",row.names(out1$summary)),])
modelSummary$parameter<-c("springOnset_spatial","springOnset_temporal","Interaction")

ggplot(modelSummary)+
  geom_crossbar(aes(x=parameter,y=mean,ymin=X2.5.,ymax=X97.5.))+
  geom_hline(yintercept=0,color="red",linetype="dashed")+
  coord_flip()+
  theme_bw()

```


