---
title: "combined_model_covariates"
author: "Diana Bowler"
date: "14 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Covariates at the level of the line

spring data (april and may)

```{r results='asis'}

setwd("/data/home/diana.bowler/climaticFiles/tempData")
library(lubridate)
library(plyr)
tempFiles <- list.files(getwd())
springData <- ldply(tempFiles, function(x){
      temp <- read.delim(x)
      temp <- subset(temp,site%in%nuIndivs$LinjeID)
      temp$Date <- gsub("temperature_","",temp$map)
      temp$Date <- as.Date(temp$Date, format="%Y_%m_%d")
      temp$day <- day(temp$Date)
      temp$month <- month(temp$Date)
      subset(temp,month%in%c(4:5))
})

save(springData,file="springData_AprilMay.RData")

```

Formatting the files for analysis

```{r,results='asis'}

setwd("/data/home/diana.bowler/climaticFiles")
load("springData_AprilMay.RData")
library(lubridate)
library(reshape2)

#get average per site/lineid
springData$temp <- as.numeric(as.character(springData$temp))
springData$Year <- year(springData$Date)
springData <- melt(tapply(springData$temp, list(springData$Year,springData$site),mean))
names(springData)<-c("Year","site","temp")

#subset to dataset
springData<-subset(springData,site%in%allData$LinjeID)

#add spatial cluster units
springData$OmradeID<-allData$OmradeID[match(springData$site,allData$LinjeID)]
springData$Fylke<-allData$Fylkesnavn[match(springData$site,allData$LinjeID)]
springData$Rapporteringsniva<-allData$Rapporteringsniva[match(springData$site,allData$LinjeID)]

#plotting
qplot(Year,temp,data=springData,geom=c("point","line"),colour=OmradeID)+
          theme(legend.position = "none")+
          facet_wrap(~Fylke)


#average across all sites
springDataSummary <- ddply(springData,.(Year),summarise,temp=mean(temp,na.rm=T))
qplot(Year,temp,data=springDataSummary,geom=c("point","line"))+theme_light()

```

#Organise the covariate

```{r, results='asis'}

#choose my covariate
myCovariate =springData
myCovariate<-subset(myCovariate,Year%in%nuIndivs$Year)
covariateMatrix = my.n
rowNames<-as.character(sapply(row.names(my.n),function(x)strsplit(as.character(x),"#")[[1]][2]))
rowNames<-as.character(sapply(rowNames,function(x)strsplit(as.character(x),"-")[[1]][2]))

all(myCovariate$site%in%rowNames)
all(myCovariate$Year%in%dimnames(covariateMatrix)[[2]])

for(i in 1:nrow(covariateMatrix)){
  for(j in 1:ncol(covariateMatrix)){
    covariateMatrix[i,j] = myCovariate$temp[myCovariate$site==rowNames[i]&
                                              myCovariate$Year==dimnames(covariateMatrix)[[2]][j]]
  }
}

#split into spatial and temporal anomalies
#spatialMatrix<-as.numeric(apply(covariateMatrix,1,function(x) median(x,na.rm=T)))
spatialMatrix<-as.numeric(rowMeans(covariateMatrix))
temporalMatrix<-as.numeric(colMeans(covariateMatrix))

#look at summaries
summary(spatialMatrix)
summary(temporalMatrix)
#more spatial than temporal variation

#make a DF
temporalDF<-data.frame(Year=dimnames(my.n)[[2]],value=temporalMatrix)
spatialDF<-data.frame(LinjeID=rowNames,value=spatialMatrix)

##make anomalies
bugs.data$temporalMatrix <- apply(covariateMatrix,2,function(x)x-spatialMatrix)
bugs.data$spatialMatrix <- t(apply(covariateMatrix,1,function(x)x-temporalMatrix))
bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix^2#also add a polynomial term for the spatial variation

#temporal
tMmelted<-melt(bugs.data$temporalMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("temporal")

#spatial
tMmelted<-melt(bugs.data$spatialMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("spatial")

```

#Alternative plotting

```{r results='asis'}

springData_Line<-ddply(springData,.(site),summarise,temp=mean(temp))
springData_Line$tempOrder <- cut(springData_Line$temp,9)
allData$tempOrder <- springData_Line$tempOrder[match(allData$LinjeID,springData_Line$site)]

#temporal
tMmelted<-melt(bugs.data$temporalMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("temporal")+
  facet_wrap(~tempOrder,scales="free")
#more positive effect in cooler areas?

#spatial
tMmelted<-melt(bugs.data$spatialMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("spatial")+
  facet_wrap(~tempOrder,scales="free")

#also plotting the rodents
allData_rodents<-ddply(allData,.(Year,Rapporteringsniva,Fylkesnavn),summarise,
                       Covariate=mean(as.numeric(SettSmagnager)/LengdeTaksert,na.rm=T),
                       totalIndiv=mean(totalIndiv/LengdeTaksert,na.rm=T))
springData_Rap<-ddply(springData,.(Rapporteringsniva),summarise,temp=mean(temp))
springData_Rap$tempOrder <- cut(springData_Rap$temp,6)
allData_rodents$tempOrder <- springData_Rap$tempOrder[match(allData_rodents$Rapporteringsniva,
                                                            springData_Rap$Rapporteringsniva)]

library(boot)
ggplot(allData_rodents,aes(x=Covariate,y=totalIndiv))+
  geom_point(aes(colour=factor(Rapporteringsniva)))+
  theme(legend.position="none")+
  ggtitle("rodents")+
  facet_wrap(~tempOrder)+
  scale_y_log10()+
  stat_smooth(method="lm")+
  facet_wrap(~Fylkesnavn)

```


Spatial pattern and temporal anomalies

```{r}

#the temporal anomalies are the same
#the spatial pattern is now assumed to be fixed
bugs.data$spatialMatrix <- spatialMatrix
bugs.data$spatialMatrix2 <-bugs.data$spatialMatrix^2

#centering
bugs.data$spatialMatrix<-bugs.data$spatialMatrix-median(bugs.data$spatialMatrix)
bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2-median(bugs.data$spatialMatrix2)

```

#specify climate as covariate 1

```{r}

bugs.data$spatialMatrix1 <- bugs.data$spatialMatrix
bugs.data$spatialMatrix1_2 <- bugs.data$spatialMatrix2
bugs.data$temporalMatrix1 <- bugs.data$temporalMatrix
summary(bugs.data$spatialMatrix1)
summary(bugs.data$temporalMatrix1)

```

#specify rodents as climate 2

#also get the rodent data - run code in the rodent Rmd

```{r}

bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix2 <- bugs.data$temporalMatrix
summary(bugs.data$spatialMatrix2)
summary(bugs.data$temporalMatrix2)

```

Centre, scale and plot

```{r}

#plot
hist(bugs.data$spatialMatrix1)
hist(as.numeric(bugs.data$temporalMatrix1))
hist(bugs.data$spatialMatrix2)
hist(as.numeric(bugs.data$temporalMatrix2))

```

Relationships between the climatic gradient and variability in the response

```{r, results='asis'}

g1<-qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix1,1,sd))+
  theme_bw()+
  scale_x_continuous(trans = "reverse")+
  xlab("Spring temperature")+ylab("Temporal variability")+ggtitle("Spring temperature")+
  stat_smooth(method="lm")

g2<-qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix2,1,sd))+
  theme_bw()+
  scale_x_continuous(trans = "reverse")+
  xlab("Spring temperature")+ylab("Temporal variability")+ggtitle("Rodents")+
  stat_smooth(method="lm")

g3<-qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix2,1,max))+
  theme_bw()+
  scale_x_continuous(trans = "reverse")+
  xlab("Spring temperature")+ylab("Maximum")+ggtitle("Rodents")+
  stat_smooth(method="lm")

g1
g2
g3

library(cowplot)
g<-plot_grid(g1, g2,labels = c("A", "B"))
#ggsave("springTemp_SD.png",width=8,height=4)

```

#model the probability to see a rodent according to climate

```{r}

#add rodent data P/A
my.rodents <- acast(allData,Line~Year,value.var="SettSmagnager",fun=max,na.rm=T)
my.rodents[is.infinite(my.rodents)]<-0
my.rodents[transectLengths==0]<-NA
all(row.names(my.n)==row.names(my.rodents))
bugs.data$rodentData <- my.rodents

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.r","line.r.sd","year.r.sd","site2.r.sd",
            "year.site2.r.sd","beta.t","beta.s","preds")

#inits <- function(){list(int.r)}

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "rodent_model.txt", n.thin=nt,
               n.chains=3, n.burnin=1000,n.iter=5000,parallel = T)

print(out1,2)

```

Plot the predicted effects of climate

```{r}
predsDF <- data.frame(out1$summary[grepl("preds",row.names(out1$summary)),])
predsDF$temp <- bugs.data$climaticGradient + median(spatialMatrix)

ggplot(predsDF)+
  geom_line(aes(x=temp,y=mean))+
  geom_ribbon(aes(x=temp,ymin=X2.5.,ymax=X97.5.),alpha=0.5)+
  theme_bw()+
  scale_x_continuous(trans = "reverse")+
  xlab("Spring mean temperature") + ylab("Probablility of rodent occurrence")

```

Model the predicted time series

```{r}

#add rodent data P/A
my.rodents <- acast(allData,Line~Year,value.var="SettSmagnager",fun=max,na.rm=T)
my.rodents[is.infinite(my.rodents)]<-0
my.rodents[transectLengths==0]<-NA
all(row.names(my.n)==row.names(my.rodents))
bugs.data$rodentData <- my.rodents

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.r","line.r.sd","year.r.sd","site2.r.sd","site.r.sd",
"year.site2.r.sd","year.site.r.sd","rodentPresence")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "rodent_preds.txt", n.thin=nt,
               n.chains=3, n.burnin=5000,n.iter=20000,parallel = T)

print(out1,2)
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_rodent_preds.RData")

```

Gets preds
```{r}
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_rodent_preds.RData")

preds<-getBUGSFits(out1,param="rodentPresence")

ggplot(preds)+
  geom_line(aes(x=Year,y=mean,colour=factor(Rapporteringsniva)))+
  facet_wrap(~Fylkesnavn)  

bugs.data$temporalMatrix2<-acast(preds,lineIndex~Year,value.var="mean")

```

get effective strip widths
```{r}
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("esw.RData")

#add to the bugs data object
bugs.data$ESW.mean <- acast(esw,lineIndex~Year,value.var="mean")
bugs.data$ESW.sd <- acast(esw,lineIndex~Year,value.var="sd")

```

#testing

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#get prioirs for density
priorDensity<-my.n*(transectLengths/1000)*(110/1000)
lineAverage<-apply(priorDensity,1,function(x)mean(x,na.rm=T))
for(i in 1:nrow(priorDensity)){
  priorDensity[i,][is.na(priorDensity[i,])]<-lineAverage[i]
}
bugs.data$priorDensity<-priorDensity

#first year
bugs.data$priorDensity1<-ceiling(as.numeric(ifelse(bugs.data$priorDensity[,1]==0,1,bugs.data$priorDensity[,1])))

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT","beta.covariateS2",
            "beta.covariate_int",
            "beta.covariate_rodT","beta.covariate_rodTL",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd","year.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")

#specific initial values
inits<- function(){
  list(predDensity=bugs.data$priorDensity)}
    
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_test.txt", n.thin=5,n.chains=3, 
             n.burnin=100,n.iter=500,parallel = T)

print(out1,2)

```

#weather interaction effects - additive effects of rodents

```{r}
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#get prioirs for density
priorDensity<-my.n*(transectLengths/1000)*(bugs.data$ESW.mean/1000)
lineAverage<-apply(priorDensity,1,function(x)mean(x,na.rm=T))
for(i in 1:nrow(priorDensity)){
  priorDensity[i,][is.na(priorDensity[i,])]<-lineAverage[i]
}
bugs.data$priorDensity<-priorDensity

#first year
bugs.data$priorDensity1<-ceiling(as.numeric(ifelse(bugs.data$priorDensity[,1]==0,1,bugs.data$priorDensity[,1])))
#bugs.data$priorDensity<-priorDensity

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT","beta.covariateS2",
            "beta.covariate_int",
            "beta.covariate_rodT","beta.covariate_rodTL",
            "obs.d.sd","line.d.sd","site2.d.sd")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")

#specific initial values
inits<- function(){
  list(beta.auto=runif(1,0,1),int.d=runif(1,0.5,3))}

#predDensity=bugs.data$priorDensity,
out1 <- jags(bugs.data, inits=inits, params, "weather_interaction_rodent_additive_esw.txt", n.thin=nt,n.chains=3, n.burnin=20000,n.iter=50000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
#save(out1,file="out1_springTemp_weather_interaction_rodent_additive.RData")
#save(out1,file="out1_springTemp_weather_interaction_rodent_additive2.RData")# second dpois term
#save(out1,file="out1_springTemp_weather_interaction_rodent_additive3.RData")# two dens dep terms
#save(out1,file="out1_springTemp_weather_interaction_rodent_additive4.RData")#quadratic spatial
#save(out1,file="out1_springTemp_weather_interaction_rodent_additive5.RData")#longer run

```

Check the model

```{r, results='asis'}

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_springTemp_weather_interaction_rodent_additive4.RData")

library(ggmcmc)

bayes.mod.fit.gg <- ggs(out1$samples,family="int.d")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateS")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateT")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)


```

plot predictions
```{r}

df<-data.frame(climaticStress=bugs.data$climaticGradient,
               predictions=out1$mean$preds,
               lower=out1$q2.5$preds,
               upper=out1$q97.5$preds)

ggplot(df)+
  geom_line(aes(x=climaticStress,y=predictions))+
  geom_ribbon(aes(x=climaticStress,ymin=lower,ymax=upper),alpha=0.5)+
  theme_bw()+
  scale_x_continuous(trans = "reverse")+
  xlab("spring temperature (spatial)")+
  ylab("effect on abundance change")+
  geom_hline(yintercept=0,colour="red",linetype="dashed")

```

interactive effects - between temporal rodent and spatial climate - incl lagged effects

```{r}
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS2_cov1",
            "beta.covariateT_cov1","beta.covariateT_cov2","beta.covariateTL_cov2",
            "beta.covariate_int","beta.covariate_intL",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd","year.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction_lagged.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=7000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springTemp_interaction_lagged.RData")

```

#Model checking

Check the model

```{r, results='asis'}

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_springTemp_interaction_lagged.RData")

library(ggmcmc)

bayes.mod.fit.gg <- ggs(out1$samples,family="int.d")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateS")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateT")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

#doesnt converge with year random effect

```

Look at coefficients

```{r}
print(out1$summary,3)

```

Plot predictions

```{r}
library(jagsUI)
out2<-update(out1,parameters.to.save=c("fit","fit.new","preds"),n.iter=2000)

bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix),max(bugs.data$spatialMatrix),length.out=1000)

#switch direction of climate

predsDF<-data.frame(Climate=bugs.data$climaticGradient,
                    Mean=out1$mean$preds,Lower=out1$q2.5$preds,Upper=out1$q97.5$preds)

ggplot(predsDF)+
  geom_line(aes(x=Climate,y=Mean))+
  geom_ribbon(aes(x=Climate,ymin=Lower,ymax=Upper))+
  theme_bw()+
  geom_hline(yintercept=0,colour="red",linetype="dashed")+
  scale_x_continuous(trans = "reverse")+
  ylab("Effect size")+
  xlab("Spring temperature")

ggsave(filename="/data/home/diana.bowler/NorskPtarmigan/plots/springtemp_rodentInt.png",
       width= 3,height=2)
  

```

#Get Bayesian p-value

```{r}
library(jagsUI)
out2<-update(out1,parameters.to.save=c("fit","fit.new"),n.iter=2000)

simslistDF<-list(fit=out2$sims.list$fit,fit.new=out2$sims.list$fit.new) 
mean(simslistDF$fit.new>simslistDF$fit) 

```

Plot predictions accounting for changes in climatic variability

```{r}

#for each climatic value, get predicted sd

#fit the model
tempdf<-data.frame(mySD=apply(bugs.data$temporalMatrix1,1,sd),
                   spatialTemp=bugs.data$spatialMatrix)
lm1<-lm(mySD~spatialTemp,data=tempdf)
#get get predictions for the model
df$predSD<-predict(lm1,newdata=list(spatialTemp=bugs.data$climaticGradient))

#multiple the effect sizes by these

ggplot(df)+
  geom_line(aes(x=climaticStress,y=predictions))+
  geom_ribbon(aes(x=climaticStress,ymin=lower,ymax=upper),alpha=0.5)+
  theme_bw()+
  scale_x_continuous(trans = "reverse")+
  xlab("spring temperature (spatial)")+
  ylab("effect on abundance change")+
  geom_hline(yintercept=0,colour="red",linetype="dashed")

```

Model with both the weather interactions and the rodent interaction

```{r}



```


