---
title: "combined_model_springOnset"
author: "Diana Bowler"
date: "14 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Extract snow depth data

```{r}
setwd("/data/home/diana.bowler/climaticFiles/snowData")
library(lubridate)

tempFiles <- list.files(getwd())
snowData <- ldply(tempFiles, function(x){
      temp <- read.delim(x)
      temp <- subset(temp,site%in%nuIndivs$LinjeID)
      temp$Date <- gsub("snow_depth_","",temp$map)
      temp$Date <- as.Date(temp$Date, format="%Y_%m_%d")
      temp$month <- month(temp$Date)
      temp$week <- week(temp$Date)
      temp <- subset(temp,month%in%c(3:5))
      temp$depth <- as.numeric(as.character(temp$depth))
      return(temp)
})

save(snowData,file="snowData_springRData")

```


#get the data on spring

```{r results='asis'}
setwd("/data/home/diana.bowler/climaticFiles")
library(lubridate)

load("snowData_spring.RData")
snowData<-subset(snowData,site %in% allData$LinjeID)
snowData<-subset(snowData,month %in% 3:4)#just use march and april
snowData$depth <- as.numeric(as.character(snowData$depth))
snowData$Year <- year(snowData$Date)

#get average per site/lineid
snowData <- tapply(snowData$depth,list(snowData$Year,snowData$site),mean)
snowData<-melt(snowData)
names(snowData)<-c("Year","site","depth")

#add further site info
snowData$OmradeID<-allData$OmradeID[match(snowData$site,allData$LinjeID)]
snowData$Fylkesnavn<-allData$Fylkesnavn[match(snowData$site,allData$LinjeID)]
snowData$Rapporteringsniva<-allData$Rapporteringsniva[match(snowData$site,allData$LinjeID)]

#plotting
ggplot(data=snowData,aes(x=Year,y=depth,group=OmradeID))+
  geom_point(aes(colour=OmradeID))+
  geom_line(aes(colour=OmradeID))+
  theme(legend.position = "none")+
  facet_wrap(~Fylkesnavn)

```

#Organise the data

```{r}
myCovariate = snowData
myCovariate = subset(myCovariate,site%in%nuIndivs$LinjeID & Year%in%nuIndivs$Year)
covariateMatrix = my.n
rowNames<-as.character(sapply(row.names(my.n),function(x)strsplit(as.character(x),"#")[[1]][2]))
rowNames<-as.character(sapply(rowNames,function(x)strsplit(as.character(x),"-")[[1]][2]))

all(rowNames%in%myCovariate$site)
all(myCovariate$Year%in%dimnames(covariateMatrix)[[2]])

for(i in 1:nrow(covariateMatrix)){
  for(j in 1:ncol(covariateMatrix)){
    covariateMatrix[i,j] = myCovariate$depth[myCovariate$site==rowNames[i]&
                                              myCovariate$Year==dimnames(covariateMatrix)[[2]][j]]
  }
}

#split into spatial and temporal anomalies
#spatialMatrix<-as.numeric(rowMeans(covariateMatrix))
spatialMatrix<-as.numeric(apply(covariateMatrix,1,function(x)median(x,na.rm=T)))
temporalMatrix<-as.numeric(colMeans(covariateMatrix))

#make a DF
temporalDF<-data.frame(Year=dimnames(my.n)[[2]],value=temporalMatrix)
spatialDF<-data.frame(OmradeID=rowNames,value=spatialMatrix)

#look at summaries
summary(spatialMatrix)
summary(temporalMatrix)

#make anomalies
bugs.data$temporalMatrix <- apply(covariateMatrix,2,function(x)x-spatialMatrix)
bugs.data$spatialMatrix <- t(apply(covariateMatrix,1,function(x)x-temporalMatrix))
bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix^2#also add a polynomial term for the spatial variation

#plot the data

#temporal
tMmelted<-melt(bugs.data$temporalMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("temporal")

#spatial
tMmelted<-melt(bugs.data$spatialMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("spatial")

```

#Alternative plotting

```{r results='asis'}

#order lines by average temperature
snowData_Line<-ddply(snowData,.(site),summarise,depth=mean(depth))
snowData_Line$depthOrder <- cut(log(snowData_Line$depth+1),9)
allData$depthOrder <- snowData_Line$depthOrder[match(allData$LinjeID,snowData_Line$site)]

#temporal
tMmelted<-melt(bugs.data$temporalMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("temporal")+
  facet_wrap(~depthOrder,scales="free")

#spatial
tMmelted<-melt(bugs.data$spatialMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("spatial")+
  facet_wrap(~depthOrder)


#also plotting the rodents
allData_rodents<-ddply(allData,.(Year,Rapporteringsniva),summarise,
                       Covariate=mean(as.numeric(SettSmagnager)/LengdeTaksert,na.rm=T),
                       totalIndiv=mean(totalIndiv/LengdeTaksert,na.rm=T))
snowData_Rap<-ddply(snowData,.(Rapporteringsniva),summarise,depth=mean(depth))
snowData_Rap$depthOrder <- cut(log(snowData_Rap$depth+1),3)
allData_rodents$depthOrder <- snowData_Rap$depthOrder[match(allData_rodents$Rapporteringsniva,snowData_Rap$Rapporteringsniva)]

library(boot)
ggplot(allData_rodents,aes(x=Covariate,y=totalIndiv))+
  geom_point(aes(colour=factor(Rapporteringsniva)))+
  theme(legend.position="none")+
  ggtitle("rodents")+
  facet_wrap(~depthOrder)+
  scale_y_log10()+
  stat_smooth(method="lm")

```

#Spatial pattern and temporal anomalies

the temporal anomalies are the same as above

the spatial pattern is now assumed to be fixed across years

```{r}

bugs.data$spatialMatrix <-spatialMatrix
bugs.data$spatialMatrix2 <-bugs.data$spatialMatrix^2
bugs.data$spatialMatrix<-bugs.data$spatialMatrix-median(bugs.data$spatialMatrix)
bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2-median(bugs.data$spatialMatrix2)

plot(bugs.data$spatialMatrix,
     apply(bugs.data$temporalMatrix,1,sd),main="spring snow depth")

#sd of variation
cor.test(bugs.data$spatialMatrix,apply(bugs.data$temporalMatrix,1,sd))#0.6576918

```

Fit glm

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","site.d.sd",
            "b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_glm.txt", n.thin=nt,
               n.chains=3, n.burnin=1000,n.iter=5000,parallel = T)

print(out1,2)

``` 

Fit glm scaled

```{r}
bugs.data$spatialMatrix<-bugs.data$spatialMatrix/sd(bugs.data$spatialMatrix)
bugs.data$temporalMatrix<-bugs.data$temporalMatrix/sd(bugs.data$temporalMatrix)

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","site.d.sd",
            "b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_glm.txt", n.thin=nt,
               n.chains=3, n.burnin=500,n.iter=2000,parallel = T)
print(out1,2)

```

#Fit a basic model against abundance

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT","predESW")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=10000,parallel = T)

save(out1,file="out1_abundance_springsnowDepth.RData")
```

#negative effects of snow depth spatially but not temporally (tends to be negative)

#plotting the results

```{r}

#effect of space
library(reshape2)
dataspatialMelt<-melt(out1$mean$pred.Space)
dataspatialMelt2<-melt(out1$q2.5$pred.Space)
dataspatialMelt3<-melt(out1$q97.5$pred.Space)
spatialMatrixMelt<-melt(bugs.data$spatialMatrix)
dataspatialMelt$springOnset<-spatialMatrixMelt$value
dataspatialMelt$lower<-dataspatialMelt2$value
dataspatialMelt$upper<-dataspatialMelt3$value

ggplot(dataspatialMelt)+geom_point(aes(x=springOnset,y=value))+geom_ribbon(aes(x=springOnset,ymin=lower,ymax=upper),alpha=0.5)+
  theme_classic()+ggtitle("Spatial variation")

#effect of time
library(reshape2)
dataspatialMelt<-melt(out1$mean$pred.Time)
dataspatialMelt2<-melt(out1$q2.5$pred.Time)
dataspatialMelt3<-melt(out1$q97.5$pred.Time)
spatialMatrixMelt<-melt(bugs.data$temporalMatrix)
dataspatialMelt$springOnset<-spatialMatrixMelt$value
dataspatialMelt$lower<-dataspatialMelt2$value
dataspatialMelt$upper<-dataspatialMelt3$value

ggplot(dataspatialMelt)+geom_point(aes(x=springOnset,y=value))+geom_ribbon(aes(x=springOnset,ymin=lower,ymax=upper),alpha=0.5)+
  theme_classic()+ggtitle("Temporal variation")


```

#Fit a model with ar1 term

```{r}

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto","beta.covariateS","beta.covariateT","pred.Time","pred.Space")

out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1.txt", n.thin=nt,
               n.chains=3, n.burnin=400,n.iter=5000,parallel = T)

print(out1,2)

save(out1,file="out1_springsnowDepth.RData")

```


Weather interaction

original units

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix),max(bugs.data$spatialMatrix),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT",
            "beta.covariate_int",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size",
            "preds")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_weather_interaction.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springsnowDepth_weather_interaction.RData")

```


#Model checking

Check the model

```{r, results='asis'}

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_springsnowDepth_weather_interaction.RData")

library(ggmcmc)

bayes.mod.fit.gg <- ggs(out1$samples,family="int.d")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateS")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateT")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

```

Look at coefficients

```{r}
print(out1$summary,3)

```

Plot predictions

```{r}
library(jagsUI)

bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix),max(bugs.data$spatialMatrix),length.out=1000)

#switch direction of climate

predsDF<-data.frame(Climate=bugs.data$climaticGradient,
                    Mean=out1$mean$preds,Lower=out1$q2.5$preds,Upper=out1$q97.5$preds)

ggplot(predsDF)+
  geom_line(aes(x=Climate,y=Mean))+
  geom_ribbon(aes(x=Climate,ymin=Lower,ymax=Upper))+
  theme_bw()+
  geom_hline(yintercept=0,colour="red",linetype="dashed")+
  #scale_x_continuous(trans = "reverse")+
  ylab("Effect size")+
  xlab("Snow depth")

ggsave(filename="/data/home/diana.bowler/NorskPtarmigan/plots/springsnowdepth_weatherInt.png",
       width= 3,height=2)
  
```

#Get Bayesian p-value

```{r}
library(jagsUI)
out2<-update(out1,parameters.to.save=c("fit","fit.new"),n.iter=2000)
simslistDF<-list(fit=out2$sims.list$fit,fit.new=out2$sims.list$fit.new) 
mean(simslistDF$fit.new>simslistDF$fit) 

```


#specify climate as covariate 1

```{r}

bugs.data$spatialMatrix1 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix1 <- bugs.data$temporalMatrix

```

#specify rodents as climate 2

#also get the rodent data - run code in the rodent Rmd

```{r}

bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix2 <- bugs.data$temporalMatrix

qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix2,1,sd),main="Rodent sd against spring onset")+stat_smooth(method="lm")

```

```{r}

#scale
#bugs.data$spatialMatrix1<-bugs.data$spatialMatrix1/sd(bugs.data$spatialMatrix1)
#bugs.data$temporalMatrix1<-bugs.data$temporalMatrix1/sd(bugs.data$temporalMatrix1)
#bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2/sd(bugs.data$spatialMatrix2)
#bugs.data$temporalMatrix2<-bugs.data$temporalMatrix2/sd(bugs.data$temporalMatrix2)

#plot
hist(bugs.data$spatialMatrix1)
hist(as.numeric(bugs.data$temporalMatrix1))
hist(bugs.data$spatialMatrix2)
hist(as.numeric(bugs.data$temporalMatrix2))

```


Relationships between the climatic gradient and variability in the response
```{r, results='asis'}

g1<-qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix1,1,sd))+
  theme_bw()+
  xlab("Spring snow depth")+ylab("Temporal variability")+ggtitle("Spring snow depth")+
  stat_smooth(method="lm")

g2<-qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix2,1,sd))+
  theme_bw()+
  xlab("Spring snow depth")+ylab("Temporal variability")+ggtitle("Rodents")+
  stat_smooth(method="lm")

g3<-qplot(bugs.data$spatialMatrix1,apply(bugs.data$temporalMatrix2,1,max))+
  theme_bw()+
  xlab("Spring snow depth")+ylab("Maximum")+ggtitle("Rodents")+
  stat_smooth(method="lm")

g4<-qplot(bugs.data$spatialMatrix1,
          apply(bugs.data$temporalMatrix1,1,sd)/abs(bugs.data$spatialMatrix1))+
  theme_bw()+
  xlab("Spring snow depth")+ylab("Temporal variability (cov)")+ggtitle("Spring snow depth")+
  stat_smooth(method="lm")

g1
g2
g3
g4

library(cowplot)
g<-plot_grid(g1, g2,labels = c("A", "B"))
ggsave("springsnowDepth_SD.png",width=8,height=4)

```


weather interaction effects -additive effects of rodents
```{r}
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT",
            "beta.covariate_int",
            "beta.covariate_rodT","beta.covariate_rodTL",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size",
            "preds")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_weather_interaction_rodent_additive.txt", n.thin=nt,
               n.chains=3, n.burnin=5000,n.iter=20000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springsnowDepth_weather_interaction_rodent_additive.RData")

```

interactive effects - between temporal rodent and spatial climate

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2",
            "beta.covariate_int",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springsnowDepth_interaction.RData")

```

Assess model fit - look at intercept, line and autocorrelation effect

```{r, results='asis'}
library(ggmcmc)
ggs_traceplot(ggs(out1$samples,family="int.d"))
ggs_traceplot(ggs(out1$samples,family="beta.auto"))
ggs_traceplot(ggs(out1$samples,family="line.d.sd"))

```

Look at coeffs of interest
```{r, results='asis'}
library(ggmcmc)
ggs_traceplot(ggs(out1$samples,family="beta.covariateS_cov1"))
ggs_traceplot(ggs(out1$samples,family="beta.covariateS_cov2"))
ggs_traceplot(ggs(out1$samples,family="beta.covariateT_cov1"))
ggs_traceplot(ggs(out1$samples,family="beta.covariateT_cov2"))
ggs_traceplot(ggs(out1$samples,family="beta.covariate_int"))

```

```{r, results='asis'}

library(ggmcmc)
ggs_density(ggs(out1$samples,family="beta.covariateS_cov1"))
ggs_density(ggs(out1$samples,family="beta.covariateS_cov2"))
ggs_density(ggs(out1$samples,family="beta.covariateT_cov1"))
ggs_density(ggs(out1$samples,family="beta.covariateT_cov2"))
ggs_density(ggs(out1$samples,family="beta.covariate_int"))

```

Plot the coefficients

```{r}

modelSummary<-data.frame(out1$summary[grepl("beta.covariate",row.names(out1$summary)),])
modelSummary$parameter<-c("springOnset_spatial","rodents_spatial","springOnset_temporal","rodents_temporal","spatialSpring x temporalRodents")

ggplot(modelSummary)+
  geom_crossbar(aes(x=parameter,y=mean,ymin=X2.5.,ymax=X97.5.))+
  geom_hline(yintercept=0,color="red",linetype="dashed")+
  coord_flip()+
  theme_bw()

```

Look at predictions

```{r}

out2<-update(out1,parameters.to.save="expNuIndivs.new",n.iter=2000)
ggs_density(ggs(out2$samples,family="expNuIndivs.new"))

my_ggs_ppmean(ggs(out1$samples,family="expNuIndivs.new"),outcome=as.numeric(bugs.data$NuIndivs))

```

Dynamics predicted by the model

```{r}
out2<-update(out1,parameters.to.save="Density",n.iter=2000)

modelSummary<-getBUGSFits(out2)

ggplot(subset(modelSummary,Year>2007),
       aes(x=Year,y=mean,group=Line))+
  geom_point(aes(colour=Kommunenavn))+
  geom_line(aes(colour=Kommunenavn))+
  facet_wrap(~Fylkesnavn)+
  theme(legend.position="none")+
  theme_bw()

```


interactive effects - between temporal rodent and spatial climate - incl lagged effects

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)


params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2","beta.covariateTL_cov2",
            "beta.covariate_int","beta.covariate_intL",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction_lagged.txt", n.thin=nt,
               n.chains=3, n.burnin=10000,n.iter=50000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_springsnowDepth_interaction_lagged.RData")

```

#Model checking

Check the model

```{r, results='asis'}

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
load("out1_springsnowDepth_interaction_lagged.RData")

library(ggmcmc)

bayes.mod.fit.gg <- ggs(out1$samples,family="int.d")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateS")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateT")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)

```

Plot predictions

```{r}
library(jagsUI)
out1<-update(out1,parameters.to.save=c("preds"),n.iter=2000)

bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix),max(bugs.data$spatialMatrix),length.out=1000)

#switch direction of climate

predsDF<-data.frame(Climate=bugs.data$climaticGradient,
                    Mean=out1$mean$preds,Lower=out1$q2.5$preds,Upper=out1$q97.5$preds)

ggplot(predsDF)+
  geom_line(aes(x=Climate,y=Mean))+
  geom_ribbon(aes(x=Climate,ymin=Lower,ymax=Upper))+
  theme_bw()+
  geom_hline(yintercept=0,colour="red",linetype="dashed")+
  #scale_x_continuous(trans = "reverse")+
  ylab("Effect size")+
  xlab("Snow depth")

ggsave(filename="/data/home/diana.bowler/NorskPtarmigan/plots/springsnowdepth_rodentInt.png",
       width= 3,height=2)
```


#compare predictions
