---
title: "combined_model_covariates"
author: "Diana Bowler"
date: "14 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

#Covariates at the level of the line

winter data

```{r results='asis'}

setwd("/data/home/diana.bowler/climaticFiles/snowData")
library(lubridate)

tempFiles <- list.files(getwd())
snowData <- ldply(tempFiles, function(x){
      temp <- read.delim(x)
      temp <- subset(temp,site%in%nuIndivs$LinjeID)
      temp$Date <- gsub("snow_depth_","",temp$map)
      temp$Date <- as.Date(temp$Date, format="%Y_%m_%d")
      temp$month <- month(temp$Date)
      temp$week <- week(temp$Date)
      temp <- subset(temp,month%in%c(8:12))
      temp$depth <- as.numeric(as.character(temp$depth))
      return(temp)
})

save(snowData,file="snowData.RData")

```

format the file

```{r}
setwd("/data/home/diana.bowler/climaticFiles")
load("snowData.RData")

#get average depth per site
outS<-tapply(snowData$depth,list(snowData$site,snowData$Date),median)
library(reshape2)
snowData<-melt(outS)
names(snowData)<-c("site","Date","depth")

#get 7 day rolling ahead means
library(zoo)
library(lubridate)
snowData$yday<-yday(as.Date(snowData$Date))
snowData$Year<-year(as.Date(snowData$Date))
snowData<-ddply(snowData,.(site,Year),function(x){
  x<-x[order(x$yday),]
  x$rollingDepth<-rollmedian(x$depth,k=7,fill=NA)
  return(x)
})
snowData<-ddply(snowData,.(site,Year),summarise,
                onset=min(yday[rollingDepth>0&!is.na(rollingDepth)]))
snowData<-subset(snowData,Year!=2017)
summary(snowData$onset)
hist(snowData$onset)

#add spatial cluster units
snowData$OmradeID<-allData$OmradeID[match(snowData$site,allData$LinjeID)]
snowData$Fylke<-allData$Fylkesnavn[match(snowData$site,allData$LinjeID)]

head(snowData)

```

#plotting

```{R}
library(ggplot2)
<<<<<<< HEAD
snowData$Fylke<-factor(snowData$Fylke,levels=FylkeOrder)

qplot(Year,onset,data=snowData,geom=c("point","line"),colour=OmradeID)+
          theme_bw()+
          theme(legend.position = "none")+
          facet_wrap(~Fylke,ncol=2)

```

Organise by along the climate stress gradient - by fylke

```{r}

#get average per fylke and organise by that
snowData_Fylke<-ddply(snowData,.(Fylke),summarise,onset=median(onset))
snowData$Fylke<-factor(snowData$Fylke,levels=snowData_Fylke$Fylke[order(snowData_Fylke$onset)])

#draw time series for each line, coloured by rapport
qplot(Year,onset,data=subset(snowData,Year>2006),geom=c("point","line"),group=site,colour=factor(OmradeID))+
          theme_bw()+
          theme(legend.position = "none")+
          facet_wrap(~Fylke,ncol=1)
=======

qplot(Year,onset,data=snowData,geom=c("point","line"),colour=OmradeID)+
          theme(legend.position = "none")+
          facet_wrap(~Fylke)

#better to calculate 7 day rolling means of julian day
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893

```


#Organise the covariate

```{r}
#choose my covariate
myCovariate =snowData
myCovariate$Year = myCovariate$Year + 1
myCovariate<-subset(myCovariate,Year%in%nuIndivs$Year)
covariateMatrix = my.n
rowNames<-as.character(sapply(row.names(my.n),function(x)strsplit(as.character(x),"#")[[1]][2]))
rowNames<-as.character(sapply(rowNames,function(x)strsplit(as.character(x),"-")[[1]][2]))

all(myCovariate$site%in%rowNames)
all(myCovariate$Year%in%dimnames(covariateMatrix)[[2]])

for(i in 1:nrow(covariateMatrix)){
  for(j in 1:ncol(covariateMatrix)){
    covariateMatrix[i,j] = myCovariate$onset[myCovariate$site==rowNames[i]&                                              myCovariate$Year==dimnames(covariateMatrix)[[2]][j]]
  }
}

#split into spatial and temporal anomalies
spatialMatrix<-as.numeric(rowMeans(covariateMatrix))
<<<<<<< HEAD
#across the 10 years of the study. Should i be longer??????
=======
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
temporalMatrix<-as.numeric(colMeans(covariateMatrix))

#look at summaries
summary(spatialMatrix)
summary(temporalMatrix)

#make a DF
temporalDF<-data.frame(Year=dimnames(my.n)[[2]],value=temporalMatrix)
spatialDF<-data.frame(LinjeID=rowNames,value=spatialMatrix)

##make anomalies
bugs.data$temporalMatrix <- apply(covariateMatrix,2,function(x)x-spatialMatrix)
bugs.data$spatialMatrix <- t(apply(covariateMatrix,1,function(x)x-temporalMatrix))
bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix^2#also add a polynomial term for the spatial variation

#temporal
tMmelted<-melt(bugs.data$temporalMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("temporal")

#spatial
tMmelted<-melt(bugs.data$spatialMatrix)
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
  geom_point(aes(colour=factor(Fylkesnavn)))+
  geom_line(aes(colour=factor(Fylkesnavn)))+
  theme(legend.position="none")+
  ggtitle("spatial")

<<<<<<< HEAD



=======
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
```

#Alternative plotting

```{r results='asis'}
tMmelted<-melt(bugs.data$temporalMatrix)
<<<<<<< HEAD
nuIndivs$CovariateT<-tMmelted$value[match(interaction(nuIndivs$Year,nuIndivs$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(nuIndivs,aes(x=CovariateT,y=nuTotal,group=LinjeID))+
=======
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("temporal")+
  facet_wrap(~Fylkesnavn,scales="free")


tMmelted<-melt(bugs.data$spatialMatrix)
<<<<<<< HEAD
nuIndivs$CovariateS<-tMmelted$value[match(interaction(nuIndivs$Year,nuIndivs$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(nuIndivs,aes(x=CovariateS,y=nuTotal,group=LinjeID))+
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("spatial")+
  facet_wrap(~Fylkesnavn)

#fixed spatial pattern
nuIndivs$CovariateSF<-spatialDF$value[match(nuIndivs$LinjeID,spatialDF$LinjeID)]

ggplot(nuIndivs,aes(x=CovariateSF,y=nuTotal,group=LinjeID))+
=======
allData$Covariate<-tMmelted$value[match(interaction(allData$Year,allData$Line),
                                           interaction(tMmelted[,2],tMmelted[,1]))]

ggplot(allData,aes(x=Covariate,y=totalIndiv,group=LinjeID))+
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
  geom_point(aes(colour=factor(OmradeID)))+
  theme(legend.position="none")+
  ggtitle("spatial")+
  facet_wrap(~Fylkesnavn)

<<<<<<< HEAD

```

Spatial pattern and temporal anomalies

```{r, results='asis'}

#the temporal anomalies are the same
#the spatial pattern is now assumed to be fixed
bugs.data$spatialMatrix <- spatialMatrix
bugs.data$spatialMatrix2 <-bugs.data$spatialMatrix^2
bugs.data$spatialMatrix<-bugs.data$spatialMatrix-mean(bugs.data$spatialMatrix)
bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2-mean(bugs.data$spatialMatrix2)

#relationship between sd of temporal effects and spatial effects
plot(bugs.data$spatialMatrix,apply(bugs.data$temporalMatrix,1,sd))
#later onser = more variable dynamics

```

What does a naive analysis say

```{r}

glm1<-glm(nuTotal~CovariateT + CovariateS, family="poisson",data=nuIndivs,offset=log(maxTransectLength))
summary(glm1)

glm1<-glm(nuTotal~CovariateT + CovariateSF, family="poisson",data=nuIndivs,offset=log(maxTransectLength))
summary(glm1)

```


Fit glm

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","site.d.sd",
            "b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_glm.txt", n.thin=nt,
               n.chains=3, n.burnin=1000,n.iter=5000,parallel = T)

```

Print output
```{r}
print(out1,2)
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_glm.RData")

```

Scaled

```{r}
bugs.data$spatialMatrix<-scale(bugs.data$spatialMatrix)

bugs.data$temporalMatrix<-scale(bugs.data$temporalMatrix)

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","site.d.sd",
            "b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_glm.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=5000,parallel = T)
print(out1,2)
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_glm_scaled.RData")
```

=======
```


Spatial pattern and temporal anomalies

```{r}

#the temporal anomalies are the same
#the spatial pattern is now assumed to be fixed
bugs.data$spatialMatrix <- cbind(spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix,spatialMatrix)
bugs.data$spatialMatrix2 <-bugs.data$spatialMatrix^2
bugs.data$spatialMatrix<-bugs.data$spatialMatrix-median(bugs.data$spatialMatrix)
bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2-median(bugs.data$spatialMatrix2)

```

>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893

#Fit a basic model against abundance

```{r}
<<<<<<< HEAD
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

params <- c("int.d","obs.d.sd","line.d.sd","b.df.0","b.group.size",
            "beta.covariateS","beta.covariateT","predESW")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=10000,parallel = T)
=======

params <- c("int.d","beta.covariateS","beta.covariateT","pred.Time","pred.Space")

out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS.txt", n.thin=nt,
               n.chains=3, n.burnin=400,n.iter=5000,parallel = T)
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893

```

#Print the output

```{r}
print(out1,2)
<<<<<<< HEAD
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_basic.RData")
#hist(out1$mean$expNuIndivs.new)
#hist(out1$mean$predESW)#varies alot...
summary(out1$mean$predESW)
=======
save(out1,file="out1_winterOnset.RData")
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893

```

#Fit a model with ar1 term
```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

<<<<<<< HEAD
params <- c("int.d","obs.d.sd","beta.auto","beta.covariateS","beta.covariateT","pred.Time","pred.Space")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1.txt", n.thin=nt,
               n.chains=3, n.burnin=10000,n.iter=30000,parallel = T)
=======
params <- c("int.d","beta.auto","beta.covariateS","beta.covariateT","pred.Time","pred.Space")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=15000,parallel = T)
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893

```

#Print the output

```{r}
print(out1,2)
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset.RData")

```

#including also the random effects

```{r}
<<<<<<< HEAD
=======

>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1
<<<<<<< HEAD
#bugs.data$TransectLength[bugs.data$TransectLength[,1]==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1_re.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=7000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_re.RData")

```

=======

params <- c("line.d.sd","site.d.sd","site2.d.sd","year.d.sd",
            "int.d","beta.auto","beta.covariateS","beta.covariateT","pred.Time","pred.Space")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1_re.txt", n.thin=nt,n.chains=3, n.burnin=2000,n.iter=10000,parallel = T)

```
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
#Print the output

```{r}
print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_re.RData")

<<<<<<< HEAD
```

Model checking

```{r}

#http://xavier-fim.net/packages/ggmcmc/
library(ggmcmc)
bayes.mod.fit.gg <- ggs(out1$samples,family="int.d")
bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateS")
bayes.mod.fit.gg <- ggs(out1$samples,family="beta.covariateT")
ggs_density(bayes.mod.fit.gg)
ggs_histogram(bayes.mod.fit.gg)
ggs_traceplot(bayes.mod.fit.gg)
ggs_running(bayes.mod.fit.gg)
ggs_compare_partial(bayes.mod.fit.gg)
ggs_autocorrelation(bayes.mod.fit.gg)
ggmcmc(bayes.mod.fit.gg, file = "bayes_fit_ggmcmc.pdf")

```

Add predictions to the data

```{r}

out2<-update(out1,parameters.to.save="expNuIndivs.new",n.iter=1000)
out3<-ggs(out2$samples,family="expNuIndivs.new")
my_ggs_density(out3)

my_ggs_ppmean(D=out3,outcome=as.numeric(my.n))
#data has a bigger tail than expected
    
```


scaled

```{r}
bugs.data$spatialMatrix<-bugs.data$spatialMatrix/sd(bugs.data$spatialMatrix)

bugs.data$temporalMatrix<-bugs.data$temporalMatrix/sd(bugs.data$temporalMatrix)

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1
#bugs.data$TransectLength[bugs.data$TransectLength[,1]==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1_re.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=5000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_re_scaled.RData")


```


#interative weather effects

original units

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix),max(bugs.data$spatialMatrix),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS","beta.covariateT",
            "beta.covariate_int",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size",
            "preds")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_weather_interaction.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_weather_interaction.RData")

```


#additive effects

#specify climate as covariate 1

```{r}

bugs.data$spatialMatrix1 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix1 <- bugs.data$temporalMatrix

```

#specify rodents as climate 2

#also get the rodent data - run code in the rodent Rmd

```{r}

bugs.data$spatialMatrix2 <- bugs.data$spatialMatrix
bugs.data$temporalMatrix2 <- bugs.data$temporalMatrix

```


Centre, scale and plot

```{r}

#scale
#bugs.data$spatialMatrix1<-bugs.data$spatialMatrix1/sd(bugs.data$spatialMatrix1)
#bugs.data$temporalMatrix1<-bugs.data$temporalMatrix1/sd(bugs.data$temporalMatrix1)
#bugs.data$spatialMatrix2<-bugs.data$spatialMatrix2/sd(bugs.data$spatialMatrix2)
#bugs.data$temporalMatrix2<-bugs.data$temporalMatrix2/sd(bugs.data$temporalMatrix2)

#plot
hist(bugs.data$spatialMatrix1)
hist(as.numeric(bugs.data$temporalMatrix1))
hist(bugs.data$spatialMatrix2)
hist(as.numeric(bugs.data$temporalMatrix2))

```

Fit additive model

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_additive.txt", n.thin=nt,
               n.chains=3, n.burnin=3000,n.iter=10000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_additive.RData")
```

interactive effects - between temporal rodent and spatial climate

```{r}

source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2",
            "beta.covariate_int",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction.txt", n.thin=nt,
               n.chains=3, n.burnin=10000,n.iter=50000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_interaction.RData")

```

interactive effects - between temporal rodent and spatial climate - incl lagged effects

```{r}
source('/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R')

#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

#predicted climatic range
bugs.data$climaticGradient<-seq(min(bugs.data$spatialMatrix1),max(bugs.data$spatialMatrix1),length.out=1000)
bugs.data$n.Preds<-length(bugs.data$climaticGradient)

params <- c("int.d","beta.auto",
            "beta.covariateS_cov1","beta.covariateS_cov2",
            "beta.covariateT_cov1","beta.covariateT_cov2","beta.covariateTL_cov2",
            "beta.covariate_int","beta.covariate_intL",
            "obs.d.sd","site.d.sd","line.d.sd","site2.d.sd",
            "b.df.0","b.group.size")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_interaction_lagged.txt", n.thin=nt,
               n.chains=3, n.burnin=10000,n.iter=50000,parallel = T)

print(out1,2)

setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_winterOnset_interaction_lagged.RData")

=======
>>>>>>> 10a97d73c42e80a4d80486cd20303edbf2d0b893
```