---
title: "combined_model_covariates"
author: "Diana Bowler"
date: "14 november 2017"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Get Habitat Data per line

```{r}

habitatdata <- read.delim(paste("/data/home/diana.bowler/NorskPtarmigan/additional_data_files","habitatInfo.txt",sep="/"))
table(habitatdata$HabitatClass)
#17 is the most sampled area - Fresh heather and dwarf-shrub communities

#get also the legend
habitats<- read.delim(paste("/data/home/diana.bowler/NorskPtarmigan/additional_data_files","habitatLegend.txt",
                            sep="/"))

```

Where are the ptarmigan?

```{r}

lineSummary <-
  allData%>%
  group_by(LinjeID) %>%
  summarise(nuTotal = sum(totalIndiv,na.rm=T))
#ptarmigan are present at all lines in pop data set

```

Combine this info with the habitat data

```{r}

lineSummary<-merge(lineSummary,habitatdata,by.x="LinjeID",by.y="ID",all=T)
lineSummary$Presence<-ifelse(is.na(lineSummary$nuTotal),0,1)
lineSummary<-merge(lineSummary,habitats,by.x="HabitatClass",by.y="Code",all.x=T)
lineSummary$nuTotal[is.na(lineSummary$nuTotal)]<-0
lineSummary$HabitatType<-ifelse(lineSummary$HabitatClass<9,"Forest","Open")
lineSummary$HabitatType[lineSummary$HabitatClass%in%c(22:25)]<-"Non-habitat"

```

Compare presence and absence proportions

```{r}

table(lineSummary$Presence,lineSummary$HabitatClass)

sort(table(lineSummary$Presence,lineSummary$Habitat)[2,]/
table(lineSummary$Habitat))

```

Draw a bar chart

```{r}

dataSummary<-ddply(lineSummary,.(Habitat,HabitatType),summarise,Presences=sum(Presence),Absences=length(Presence[Presence==0]))
dataSummaryMelt<-melt(dataSummary,id=c("Habitat","HabitatType"))

ggplot(dataSummaryMelt)+
  geom_bar(aes(x=Habitat,y=value,fill=variable),stat="identity")+
  ylab("number of sites")+
  coord_flip()

```


Options:
- % forest habitat
- % open/natural habitat

```{r results='asis'}

lineSummary$propOpen <- lineSummary$Open/lineSummary$TotalTerrestrial
lineSummary$propForest <- lineSummary$Forest/lineSummary$TotalTerrestrial

summary(lineSummary$propOpen)
summary(lineSummary$propForest)

par(mfrow=c(2,2))
hist(lineSummary$propOpen)
hist(asin(sqrt(lineSummary$propOpen)))
hist(lineSummary$propForest)
hist(asin(sqrt(lineSummary$propForest)))

setwd("/data/home/diana.bowler/NorskPtarmigan/additional_data_files")
save(lineSummary,file="lineSummary.RData")

```

Get variables ready for the data files

Proportion of forest is strongly negatively related to the proportion of open habitat

```{r}

#order by site info
siteInfo$propForest <- lineSummary$propForest[match(siteInfo$originalLinjeID,lineSummary$LinjeID)]
siteInfo$propOpen <- lineSummary$propOpen[match(siteInfo$originalLinjeID,lineSummary$LinjeID)]

#add to bugs data with a normalising transformation (arcsine sqrt)
bugs.data$propForest <- asin(sqrt(siteInfo$propForest))
bugs.data$propOpen <- asin(sqrt(siteInfo$propOpen))

qplot(propForest,propOpen,data=siteInfo)

```


#Fit a model (no ar1)

```{r}
source("/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R")
                                
setwd("/data/home/diana.bowler/NorskPtarmigan/models")

out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariate_habitat.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=10000,parallel = T)

```

Print the output

```{r}

print(out1,2)

save(out1,file="/data/home/diana.bowler/NorskPtarmigan/model-outputs/out1_habitat.RData")

```



#Fit a model with ar1 term

```{r}
source("/data/home/diana.bowler/NorskPtarmigan/script/bugsFunctions.R")
                                
#year one priors
bugs.data$year1<-as.numeric(apply(my.n,1,function(x)ifelse(!is.na(x[1]),x[1],max(x,na.rm=T))))
bugs.data$year1[bugs.data$year1==0]<-1

params <- c("int.d","beta.auto","beta.covariateO")

setwd("/data/home/diana.bowler/NorskPtarmigan/models")
out1 <- jags(bugs.data, inits=NULL, params, "combined_model_covariateTS_ar1_habitat.txt", n.thin=nt,
               n.chains=3, n.burnin=2000,n.iter=10000,parallel = T)

```

Print the output

```{r}

print(out1,2)
setwd("/data/home/diana.bowler/NorskPtarmigan/model_outputs")
save(out1,file="out1_habitat.RData")

```


